---
title: "Master Thesis Round Three"
author: "Patrick Sssonzi"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = FALSE)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
library(readxl)
library(dplyr)
library(readr)
library(ggplot2)
library(kableExtra)

```


# FLOW Vs SPEED + DENSITY

## Enter Data
```{r, echo=TRUE}
All_Data_new = read_excel("Final_Working_Data_File.xlsx", 
                          sheet = "Big_Data_SPSS")


```



## Test normality of data
```{r, echo=TRUE}
hist(All_Data_new$`Mean Speed (m/min)`)

qqnorm(All_Data_new$`Mean Speed (m/min)`, main='Mean Speed (m/min)')
qqline(All_Data_new$`Mean Speed (m/min)`)


```


##### --- One-sample Kolmogorov-Smirnov test
```{r, echo=FALSE}
ks.test(All_Data_new$`Mean Speed (m/min)`, 'pnorm')
ks.test(All_Data_new$`Pedestrian Density (Peds/m2)`, 'pnorm')
ks.test(All_Data_new$`Pedestrian Flow (peds/m/min)`, 'pnorm')

```


#### --- Lilliefors (Kolmogorov-Smirnov) test for normality
```{r, echo=FALSE}
nortest::lillie.test(All_Data_new$`Mean Speed (m/min)`)
nortest::lillie.test(All_Data_new$`Pedestrian Density (Peds/m2)`)
nortest::lillie.test(All_Data_new$`Pedestrian Flow (peds/m/min)`)

```


#### --- Anderson-Darling test for normality
```{r, echo=FALSE}
nortest::ad.test(All_Data_new$`Mean Speed (m/min)`)
nortest::ad.test(All_Data_new$`Pedestrian Density (Peds/m2)`)
nortest::ad.test(All_Data_new$`Pedestrian Flow (peds/m/min)`)

```


#### --- Cramer-von Mises test for normality
```{r, echo=FALSE}
nortest::cvm.test(All_Data_new$`Mean Speed (m/min)`)
nortest::cvm.test(All_Data_new$`Pedestrian Density (Peds/m2)`)
nortest::cvm.test(All_Data_new$`Pedestrian Flow (peds/m/min)`)
```


#### --- Pearson chi-square test for normality
```{r}
nortest::pearson.test(All_Data_new$`Mean Speed (m/min)`)
nortest::pearson.test(All_Data_new$`Pedestrian Density (Peds/m2)`)
nortest::pearson.test(All_Data_new$`Pedestrian Flow (peds/m/min)`)

```


\newpage

## ---- Run ANOVA or non-parametric alternative

### TWO WAY ANOVA WITHOUT INTERACTION EFFECTS
```{r, echo=TRUE}
anova_without_interactions <- aov(`Mean Speed (m/min)` ~ Location +
                                   Time + Sex, 
                                   data = All_Data_new)
summary(anova_without_interactions)

hist(residuals(anova_without_interactions))
nortest::ad.test(residuals(anova_without_interactions))

```

We can see clearly that the data is **NOT normally distributed.**


### TWO WAY ANOVA WITH INTERACTION EFFECTS
```{r, echo=TRUE}
anova_with_interactions <- aov(`Mean Speed (m/min)` ~ Location + Time 
                               + Sex + Location:Time + Location:Sex +
                                 Time:Sex, 
                                  data = All_Data_new)

summary(anova_with_interactions)

```


### ---- Check ANOVA assumptions: test validity?

#### 1. Check the homogeneity of variance assumption
```{r}
plot(anova_without_interactions, 1)

```


#### Use the Levene's test to check the homogeneity of variances.
```{r}
library(car)
car::leveneTest(`Mean Speed (m/min)` ~ Location * Time * Sex, 
                data = All_Data_new)

```


### 2. Check the normality assumption

### Normality plot of the residuals Q-Q plot
```{r}
plot(anova_without_interactions, 2)

```

### cross-check with Anderson-Darling test test about normality
```{r}
nortest::ad.test(residuals(anova_without_interactions) )

```



## Split data into training data set and validation data set
```{r}
n=dim(All_Data_new)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
All_Data_new_train = All_Data_new[id,]
All_Data_new_validation = All_Data_new[-id,]

```


## 2D graph

```{r}
ggplot(data = All_Data_new_train, 
       aes(x = `Pedestrian Density (Peds/m2)`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="All data: Pedestrain Flow Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```


## -- three dimensional models

##### with intercept included and linear model
```{r}

summary(All_Data_new_unknown <- 
          lm(data = All_Data_new_train, 
             formula = `Pedestrian Flow (peds/m/min)`  ~ 
               `Pedestrian Density (Peds/m2)` +
               `Mean Speed (m/min)` ))

```


##### with intercept removed and linear model
```{r}

summary(All_Data_new_unknown <- 
          lm(data = All_Data_new_train, 
             formula = `Pedestrian Flow (peds/m/min)`  ~ 
               0 + `Pedestrian Density (Peds/m2)` +
                   `Mean Speed (m/min)` ))

```


##### with intercept removed and equation of an ellipse
```{r}

summary(All_Data_new_unknown <- 
          lm(data = All_Data_new_train, 
             formula = `Pedestrian Flow (peds/m/min)`  ~ 
               0 + (`Pedestrian Density (Peds/m2)`)^2 +
               (`Mean Speed (m/min)`)^2 ))

```


##### with intercept removed and polynomial of degree 2
```{r}

summary(All_Data_new_unknown <- 
          lm(data = All_Data_new_train, 
             formula = `Pedestrian Flow (peds/m/min)`  ~ 
               0 + poly(`Pedestrian Density (Peds/m2)`,2, raw = T) +
               poly(`Mean Speed (m/min)`,2, raw = T)))

```


##### with intercept removed and polynomial of degree 2 and transformation
```{r}

summary(All_Data_new_unknown <- 
          lm(data = All_Data_new_train, 
          formula = `Pedestrian Flow (peds/m/min)`  ~ 
          0 + poly(`Pedestrian Density (Peds/m2)`,2, raw = T) +
          poly(1/(log10(`Mean Speed (m/min)`)),2, raw = T)))


summary(All_Data_new_unknown <- 
          lm(data = All_Data_new_train, 
          formula = `Pedestrian Flow (peds/m/min)`  ~ 
          0 + poly(`Pedestrian Density (Peds/m2)`,2, raw = T) +
          poly(1/(sqrt(`Mean Speed (m/min)`)),2, raw = T)))

summary(All_Data_new_unknown <- 
          lm(data = All_Data_new_train, 
          formula = `Pedestrian Flow (peds/m/min)`  ~ 
          0 + poly(1/(log10(`Pedestrian Density (Peds/m2)`)),2, raw = T) +
          poly(1/(log10(`Mean Speed (m/min)`)),2, raw = T)))


coefficients(All_Data_new_unknown)


```

\newpage


# Sunday March 27th 2022
```{r}

Masters_Research_Final_for_Modelling = 
  read_excel("Masters Research Final for Modelling.xls", 
              sheet = "Only_Data")

LOS_A = subset(Masters_Research_Final_for_Modelling,
               subset = (LOS == "A") )


summary(flow_model <- 
          lm(data = LOS_A, 
             formula = log(Flow)  ~ 
               0 + log(Speed) + log(Density) ) )


summary(flow_model <- 
          lm(data = LOS_A, 
             formula = log(Flow)  ~ 
               0 + log(Speed^(-1)) + log( Density^(-1) ) ) )

summary(flow_model <- 
          lm(data = LOS_A, 
             formula = log(Flow)  ~ 
               0 + log(Speed^2) + log( Density^2 ) ) )



LOS_B = subset(Masters_Research_Final_for_Modelling,
               subset = (LOS == "B") )


summary(flow_model <- 
          lm(data = LOS_B, 
             formula = log(Flow)  ~ 
               0 + log(Speed) + log(Density) ) )


summary(flow_model <- 
          lm(data = LOS_B, 
             formula = log(Flow)  ~ 
               0 + log(Speed^(-1)) + log( Density^(-1) ) ) )

summary(flow_model <- 
          lm(data = LOS_B, 
             formula = log(Flow)  ~ 
               0 + log(Speed^2) + log( Density^2 ) ) )


LOS_C = subset(Masters_Research_Final_for_Modelling,
               subset = (LOS == "C") )


summary(flow_model <- 
          lm(data = LOS_C, 
             formula = log(Flow)  ~ 
               0 + log(Speed) + log(Density) ) )


summary(flow_model <- 
          lm(data = LOS_C, 
             formula = log(Flow)  ~ 
               0 + log(Speed^(-1)) + log( Density^(-1) ) ) )

summary(flow_model <- 
          lm(data = LOS_C, 
             formula = log(Flow)  ~ 
               0 + log(Speed^2) + log( Density^2 ) ) )




LOS_D = subset(Masters_Research_Final_for_Modelling,
               subset = (LOS == "D") )


summary(flow_model <- 
          lm(data = LOS_D, 
             formula = log(Flow)  ~ 
               0 + log(Speed) + log(Density) ) )


summary(flow_model <- 
          lm(data = LOS_D, 
             formula = log(Flow)  ~ 
               0 + log(Speed^(-1)) + log( Density^(-1) ) ) )

summary(flow_model <- 
          lm(data = LOS_D, 
             formula = log(Flow)  ~ 
               0 + log(Speed^2) + log( Density^2 ) ) )



LOS_E = subset(Masters_Research_Final_for_Modelling,
               subset = (LOS == "E") )


summary(flow_model <- 
          lm(data = LOS_E, 
             formula = log(Flow)  ~ 
               0 + log(Speed) + log(Density) ) )


summary(flow_model <- 
          lm(data = LOS_E, 
             formula = log(Flow)  ~ 
               0 + log(Speed^(-1)) + log( Density^(-1) ) ) )

summary(flow_model <- 
          lm(data = LOS_E, 
             formula = log(Flow)  ~ 
               0 + log(Speed^2) + log( Density^2 ) ) )



LOS_F = subset(Masters_Research_Final_for_Modelling,
               subset = (LOS == "F") )


summary(flow_model <- 
          lm(data = LOS_F, 
             formula = log(Flow)  ~ 
               0 + log(Speed) + log(Density) ) )


summary(flow_model <- 
          lm(data = LOS_F, 
             formula = log(Flow)  ~ 
               0 + log(Speed^(-1)) + log( Density^(-1) ) ) )

summary(flow_model <- 
          lm(data = LOS_F, 
             formula = log(Flow)  ~ 
               0 + log(Speed^2) + log( Density^2 ) ) )


```
