---
title: "Master Thesis Round Two"
author: "Patrick Sssonzi"
date: "1/29/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = FALSE)
library(ggplot2)
library(readxl)
library(kableExtra)
library(plotly)
```


# SPEED Vs DENSITY

## Enter Data
```{r, echo=TRUE}
Bombo <- read_excel("Speed Density Flow.xlsx", 
                                 sheet = "Bombo-Clean")

Entebbe <- read_excel("Speed Density Flow.xlsx", 
                      sheet = "Entebbe-Clean")

Hoima <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Hoima-Clean")


Jinja <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Jinja-Clean")

```


## Split data into training data set and validation data set

```{r, echo=TRUE}
n=dim(Bombo)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Bombo_train = Bombo[id,]
Bombo_validation = Bombo[-id,]


n=dim(Entebbe)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Entebbe_train = Entebbe[id,]
Entebbe_validation = Entebbe[-id,]


n=dim(Hoima)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Hoima_train = Hoima[id,]
Hoima_validation = Hoima[-id,]


n=dim(Jinja)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Jinja_train = Jinja[id,]
Jinja_validation = Jinja[-id,]

```


## Bombo: graph and linear regression and validation
```{r, echo=FALSE}
ggplot(data = Bombo_train, 
       aes(x = `Pedestrian Density (Peds/m2) b`, y = `(Mean Speed m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = FALSE, color="black", formula = y~x)+
  labs(title="Bombo: Pedestrain Speed Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))


summary(Bombo_lm <- lm(data = Bombo_train, formula = `(Mean Speed m/min)`~ `Pedestrian Density (Peds/m2) b`))


predict_Bombo_density <- predict(Bombo_lm, 
                         newdata = Bombo_validation)

mse_bombo <- mean((predict_Bombo_density - 
        Bombo_validation$`(Mean Speed m/min)`)**2)

rmse_bombo <- sqrt(mean((predict_Bombo_density - 
             Bombo_validation$`(Mean Speed m/min)`)**2))

ggplot()+
  aes(y = predict_Bombo_density, 
      x = Bombo_validation$`(Mean Speed m/min)`) + 
  geom_point(color='red')+
  geom_smooth(method = "lm", se = TRUE, color="black", formula = y~x)+
  labs(title="Speed-Density Model Validation for Bombo road", 
       x="Actual Walking Speed m/min", y=
         "Predicted Walking Speed m/min)")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))
  

```

## Entebbe: graph and linear regression and validation

```{r, echo=FALSE}
ggplot(data = Entebbe_train, aes(x = `Pedestrian Density (Peds/m2) b`, y = `(Mean Speed m/min)`)) + 
  geom_point(color='red')+
  geom_smooth(method = "lm", se = TRUE, color="black", formula = y~x)+
  labs(title="Entebbe: Pedestrain Speed Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))

summary(Entebbe_lm <-lm(data = Entebbe_train, formula = `(Mean Speed m/min)`~ `Pedestrian Density (Peds/m2) b`))


### Entebbe error 
mean(( (Entebbe_pred<-predict(Entebbe_lm, 
              newdata = Entebbe_validation)) 
      - Entebbe_validation$`(Mean Speed m/min)`)**2)


ggplot()+
  aes(y = Entebbe_pred, 
      x = Entebbe_validation$`(Mean Speed m/min)`) + 
  geom_point(color='red')+
  geom_smooth(method = "lm", se = TRUE, color="black", formula = y~x)+
  labs(title="Speed-Density Model Validation for Entebbe road", 
       x="Actual Walking Speed m/min", 
       y="Predicted Walking Speed m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))




```

## Hoima: graph and linear regression and validation

```{r, echo=FALSE}
ggplot(data = Hoima_train, aes(x = `Pedestrian Density (Peds/m2) b`, y = `(Mean Speed m/min)`)) + 
  geom_point(color='black')+
  geom_smooth(method = "lm", se = TRUE, color="black", formula = y~x)+
  labs(title="Hoima: Pedestrain Speed Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))

summary(Hoima_lm <-lm(data = Hoima_train, formula = `(Mean Speed m/min)`~ `Pedestrian Density (Peds/m2) b`))


### Hoima error 
mean(( (Hoima_pred <- predict(Hoima_lm, 
                newdata = Hoima_validation)) 
       - Hoima_validation$`(Mean Speed m/min)`)**2, na.rm = TRUE)



ggplot()+
  aes(y = Hoima_pred, 
      x = Hoima_validation$`(Mean Speed m/min)`) + 
  geom_point(color='red')+
  geom_smooth(method = "lm", se = TRUE, color="black", formula = y~x)+
  labs(title="Speed-Density Model Validation for Hoima road", 
       x="Actual Walking Speed m/min", 
       y="Predicted Walking Speed m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))



```

## Jinja: graph and linear regression and validation

```{r, echo=FALSE}
ggplot(data = Jinja_train, aes(x = `Pedestrian Density (Peds/m2) b`, y = `(Mean Speed m/min)`)) + 
  geom_point(color='black')+
  geom_smooth(method = "lm", se = TRUE, color="black", formula = y~x)+
  labs(title="Jinja: Pedestrain Speed Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))

summary(Jinja_lm <-lm(data = Jinja_train, formula = `(Mean Speed m/min)`~ `Pedestrian Density (Peds/m2) b`))


### Jinja error 
mean(( (Jinja_pred <- predict(Jinja_lm, 
                newdata = Jinja_validation)) 
       - Jinja_validation$`(Mean Speed m/min)`)**2, na.rm = TRUE)


ggplot()+
  aes(y = Jinja_pred, 
      x = Jinja_validation$`(Mean Speed m/min)`) + 
  geom_point(color='red')+
  geom_smooth(method = "lm", se = TRUE, color="black", formula = y~x)+
  labs(title="Speed-Density Model Validation for Jinja road", 
       x="Actual Walking Speed m/min", 
       y="Predicted Walking Speed m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))

```

## Confidence intervals
```{r}

#### Bombo
confint(lm(data = Bombo_train, formula = `(Mean Speed m/min)`~ `Pedestrian Density (Peds/m2) b`), level = .95)

#### Entebbe
confint(lm(data = Entebbe_train, formula = `(Mean Speed m/min)`~ `Pedestrian Density (Peds/m2) b`))


#### Hoima
confint(lm(data = Hoima_train, formula = `(Mean Speed m/min)`~ `Pedestrian Density (Peds/m2) b`))


#### Jinja
confint(lm(data = Jinja_train, formula = `(Mean Speed m/min)`~ `Pedestrian Density (Peds/m2) b`))

```


\newpage


# FLOW Vs DENSITY


## Enter Data
```{r, echo=TRUE}
Bombo <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Bombo-Reduced")

Entebbe <- read_excel("Speed Density Flow.xlsx", 
                      sheet = "Entebbe-Reduced")

Hoima <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Hoima-Reduced")


Jinja <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Jinja-Reduced")
```

### Split data into training data set and validation data set
```{r, echo=TRUE}
Bombo[,12] <- Bombo[,12]*10
n=dim(Bombo)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Bombo_train = Bombo[id,]
Bombo_validation = Bombo[-id,]



Entebbe[,12] <- Entebbe[,12]*10
n=dim(Entebbe)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Entebbe_train = Entebbe[id,]
Entebbe_validation = Entebbe[-id,]



Hoima[,12] <- Hoima[,12]*10
n=dim(Hoima)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Hoima_train = Hoima[id,]
Hoima_validation = Hoima[-id,]



Jinja[,12] <- Jinja[,12]*10
n=dim(Jinja)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Jinja_train = Jinja[id,]
Jinja_validation = Jinja[-id,]

```


## Bombo

```{r}
ggplot(data = Bombo_train, 
       aes(x = `Pedestrian Density (Peds/m2) b`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="Bombo: Pedestrain Flow Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))
```


### Bombo training summary
```{r}
summary(Bombo_poly <- lm(data = Bombo_train, 
            formula = `Pedestrian Flow (peds/m/min)`  ~ 
            0 + poly(`Pedestrian Density (Peds/m2) b`,2, raw = T)))

coefficients(Bombo_poly)
# coef(Bombo_poly)

confint(Bombo_poly, level = 0.95)
```

### Model Validation on validation data set - MSE and RMSE
```{r}
predict_Bombo_poly <- predict(Bombo_poly, 
                        newdata = Bombo_validation)

### Model Validation Bombo

ggplot()+
  aes(x = Bombo_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Bombo_poly)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow-Density Model Validation for Bombo road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
          panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```



## Entebbe

```{r}
ggplot(data = Entebbe_train, 
       aes(x = `Pedestrian Density (Peds/m2) b`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="Entebbe: Pedestrain Flow Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))

```


### Entebbe training summary
```{r}
summary(Entebbe_poly <- lm(data = Entebbe_train, 
                         formula = `Pedestrian Flow (peds/m/min)`  ~ 
                0 + poly(`Pedestrian Density (Peds/m2) b`,2, raw = T)))

coefficients(Entebbe_poly)
# coef(Entebbe_poly)

confint(Entebbe_poly, level = 0.95)
```

### Model Validation on validation data set - MSE and RMSE
```{r}
predict_Entebbe_poly <- predict(Entebbe_poly, 
                              newdata = Entebbe_validation)

### Model Validation Entebbe

ggplot()+
  aes(x = Entebbe_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Entebbe_poly)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow-Density Model Validation for Entebbe road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```



## Hoima

```{r}
ggplot(data = Hoima_train, 
       aes(x = `Pedestrian Density (Peds/m2) b`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="Hoima: Pedestrain Flow Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))

```


### Hoima training summary
```{r}
summary(Hoima_poly <- lm(data = Hoima_train, 
                formula = `Pedestrian Flow (peds/m/min)`  ~ 
              0 + poly(`Pedestrian Density (Peds/m2) b`,2, raw = T)))

coefficients(Hoima_poly)
# coef(Hoima_poly)

confint(Hoima_poly, level = 0.95)
```

### Model Validation on validation data set - MSE and RMSE
```{r}

predict_Hoima_poly <- predict(Hoima_poly, 
                              newdata = Hoima_validation)

### Model Validation Hoima

ggplot()+
  aes(x = Hoima_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Hoima_poly)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow-Density Model Validation for Hoima road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```




## Jinja

```{r}
ggplot(data = Jinja_train, 
       aes(x = `Pedestrian Density (Peds/m2) b`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="Jinja: Pedestrain Flow Vs Pedestrian Density")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))

```


### Jinja training summary
```{r}
summary(Jinja_poly <- lm(data = Jinja_train, 
                         formula = `Pedestrian Flow (peds/m/min)`  ~ 
                0 + poly(`Pedestrian Density (Peds/m2) b`,2,raw = T)))

coefficients(Jinja_poly)

confint(Jinja_poly, level = 0.95)
```

### Model Validation on validation data set - MSE and RMSE
```{r}

predict_Jinja_poly <- predict(Jinja_poly, 
                              newdata = Jinja_validation)

### Model Validation Jinja

ggplot()+
  aes(x = Jinja_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Jinja_poly)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow-Density Model Validation for Jinja road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```


## Maximum values
```{r}

#### Bombo
#### Max Flow
max(Bombo$`Pedestrian Flow (peds/m/min)`)
#### Max Density
Bombo$`Pedestrian Density (Peds/m2) b`[
  which.max(Bombo$`Pedestrian Flow (peds/m/min)`)]


#### Entebbe
#### Max Flow
max(Entebbe$`Pedestrian Flow (peds/m/min)`)
#### Max Density
Entebbe$`Pedestrian Density (Peds/m2) b`[
  which.max(Entebbe$`Pedestrian Flow (peds/m/min)`)]


#### Hoima
#### Max Flow
max(Hoima$`Pedestrian Flow (peds/m/min)`, na.rm = TRUE)
#### Max Density
Hoima$`Pedestrian Density (Peds/m2) b`[
  which.max(Hoima$`Pedestrian Flow (peds/m/min)`)]


#### Jinja
#### Max Flow
max(Jinja$`Pedestrian Flow (peds/m/min)`, na.rm = TRUE)
#### Max Density
Jinja$`Pedestrian Density (Peds/m2) b`[
  which.max(Jinja$`Pedestrian Flow (peds/m/min)`)]


```


\newpage


# FLOW Vs SPEED

## Enter Data
```{r, echo=TRUE}
Bombo <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Bombo-Reduced")

Entebbe <- read_excel("Speed Density Flow.xlsx", 
                      sheet = "Entebbe-Reduced")

Hoima <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Hoima-Reduced")


Jinja <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Jinja-Reduced")
```



## Split data into training data set and validation data set

```{r}
n=dim(Bombo)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Bombo_train = Bombo[id,]
Bombo_validation = Bombo[-id,]


n=dim(Entebbe)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Entebbe_train = Entebbe[id,]
Entebbe_validation = Entebbe[-id,]


n=dim(Hoima)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Hoima_train = Hoima[id,]
Hoima_validation = Hoima[-id,]


n=dim(Jinja)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Jinja_train = Jinja[id,]
Jinja_validation = Jinja[-id,]

```


## Bombo

### Graph
```{r}
ggplot(data = Bombo_train, 
       aes(x = `(Mean Speed m/min)`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="Bombo: Pedestrain Flow Vs Pedestrian Speed")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```


### Model Summary
```{r}
summary(Bombo_poly <- lm(data = Bombo_train, 
      formula = `Pedestrian Flow (peds/m/min)`  ~ 
      0 + poly(`(Mean Speed m/min)`,2, raw = T)))

coefficients(Bombo_poly)
# coef(Bombo_poly)

confint(Bombo_poly, level = 0.95)

predict_Bombo_poly <- predict(Bombo_poly, 
                              newdata = Bombo_validation)

```


### Model Validation Bombo
```{r}
ggplot()+
  aes(x = Bombo_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Bombo_poly)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow-Speed Model Validation for Bombo road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```



## Entebbe

### Graph
```{r}
ggplot(data = Entebbe_train, 
       aes(x = `(Mean Speed m/min)`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="Entebbe: Pedestrain Flow Vs Pedestrian Speed")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))


```


### Model Summary
```{r}
summary(Entebbe_poly <- lm(data = Entebbe_train, 
                         formula = `Pedestrian Flow (peds/m/min)`  ~ 
                           0 + poly(`(Mean Speed m/min)`,2, raw = T)))

coefficients(Entebbe_poly)
# coef(Bombo_poly)

confint(Entebbe_poly, level = 0.95)

predict_Entebbe_poly <- predict(Entebbe_poly, 
                              newdata = Entebbe_validation)

```


### Model Validation Entebbe
```{r}
ggplot()+
  aes(x = Entebbe_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Entebbe_poly)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow-Speed Model Validation for Entebbe road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```



## Hoima

### Graph
```{r}
ggplot(data = Hoima_train, 
       aes(x = `(Mean Speed m/min)`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="Hoima: Pedestrain Flow Vs Pedestrian Speed")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))


```


### Model Summary
```{r}
summary(Hoima_poly <- lm(data = Hoima_train, 
                         formula = `Pedestrian Flow (peds/m/min)`  ~ 
                           0 + poly(`(Mean Speed m/min)`,2, raw = T)))

coefficients(Hoima_poly)
# coef(Bombo_poly)

confint(Hoima_poly, level = 0.95)

predict_Hoima_poly <- predict(Bombo_poly, 
                              newdata = Hoima_validation)

```


### Model Validation Hoima
```{r}
ggplot()+
  aes(x = Hoima_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Hoima_poly)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow-Speed Model Validation for Hoima road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))
```



## Jinja

### Graph
```{r}
ggplot(data = Jinja_train, 
       aes(x = `(Mean Speed m/min)`, 
           y = `Pedestrian Flow (peds/m/min)`)) + 
  geom_point(color='red') +
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y ~ poly(x, 2, raw = T))+
  labs(title="Jinja: Pedestrain Flow Vs Pedestrian Speed")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.x = element_text(size = 8))+
  theme(axis.title.y = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```


### Model Summary
```{r}
summary(Jinja_poly <- lm(data = Jinja_train, 
                         formula = `Pedestrian Flow (peds/m/min)`  ~ 
                           0 + poly(`(Mean Speed m/min)`,2, raw = T)))

coefficients(Jinja_poly)
# coef(Bombo_poly)

confint(Jinja_poly, level = 0.95)

predict_Jinja_poly <- predict(Jinja_poly, 
                              newdata = Jinja_validation)

```


### Model Validation Jinja
```{r}
ggplot()+
  aes(x = Jinja_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Jinja_poly)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow-Speed Model Validation for Jinja road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))
```



\newpage


# FLOW Vs DENSITY + SPEED (THREE DIMENSIONAL)

## Enter Data
```{r, echo=TRUE}
Bombo <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Bombo-Reduced")

Entebbe <- read_excel("Speed Density Flow.xlsx", 
                      sheet = "Entebbe-Reduced")

Hoima <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Hoima-Reduced")


Jinja <- read_excel("Speed Density Flow.xlsx", 
                    sheet = "Jinja-Reduced")


Combined_2 = rbind(Bombo,Entebbe,Hoima,Jinja)
Combined_2[,12] <- Combined_2[,12]*10

```



## Split data into training data set and validation data set

```{r}
n=dim(Bombo)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Bombo_train = Bombo[id,]
Bombo_validation = Bombo[-id,]


n=dim(Entebbe)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Entebbe_train = Entebbe[id,]
Entebbe_validation = Entebbe[-id,]


n=dim(Hoima)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Hoima_train = Hoima[id,]
Hoima_validation = Hoima[-id,]


n=dim(Jinja)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Jinja_train = Jinja[id,]
Jinja_validation = Jinja[-id,]

```


## Bombo

### Model Summary three dimension
```{r}
summary(Bombo_unknown <- lm(data = Bombo_train, 
                         formula = `Pedestrian Flow (peds/m/min)`  ~ 
                          0 + poly(`Pedestrian Density (Peds/m2) b`,2, raw = T) +
                          poly(1/(log10(`(Mean Speed m/min)`)),2, raw = T)))

coefficients(Bombo_unknown)

```


### Model Validation Bombo three dimensions
```{r}

predict_Bombo_poly_3 <- predict(Bombo_unknown, 
                                  newdata = Bombo_validation)

ggplot()+
  aes(x = Bombo_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Bombo_poly_3)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow = Density + Speed : Model Validation for Bombo road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```



## Entebbe


### Model Summary
```{r}
summary(Entebbe_unknown <- lm(data = Entebbe_train, 
                            formula = `Pedestrian Flow (peds/m/min)`  ~ 
                              0 + poly(`Pedestrian Density (Peds/m2) b`,2, raw = T) +
                              poly(1/(log10(`(Mean Speed m/min)`)),2, raw = T)))


coefficients(Entebbe_unknown)

```


### Model Validation Entebbe three dimensions
```{r}
predict_Entebbe_poly_3 <- predict(Entebbe_unknown, 
                                newdata = Entebbe_validation)

ggplot()+
  aes(x = Entebbe_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Entebbe_poly_3)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow = Density + Speed : Model Validation for Entebbe road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```



## Hoima


### Model Summary three dimensions
```{r}
summary(Hoima_unknown <- lm(data = Hoima_train, 
                            formula = `Pedestrian Flow (peds/m/min)`  ~ 
                            0 + poly(`Pedestrian Density (Peds/m2) b`,2, raw = T) +
                            poly(1/(log10(`(Mean Speed m/min)`)),2, raw = T)))

coefficients(Hoima_unknown)

```


### Model Validation Hoima three dimensions
```{r}
predict_Hoima_poly_3 <- predict(Hoima_unknown, 
                                newdata = Hoima_validation)

ggplot()+
  aes(x = Hoima_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Hoima_poly_3)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow = Density + Speed : Model Validation for Hoima road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

```



## Jinja


### Model Summary three dimensions
```{r}
summary(Jinja_unknown <- lm(data = Jinja_train, 
                            formula = `Pedestrian Flow (peds/m/min)`  ~ 
                              0 + poly(`Pedestrian Density (Peds/m2) b`,2, raw = T) +
                              poly(1/(log10(`(Mean Speed m/min)`)),2, raw = T)))

coefficients(Jinja_unknown)

```


### Model Validation Jinja three dimensions
```{r}

predict_Jinja_poly_3 <- predict(Jinja_unknown, 
                                newdata = Jinja_validation)

ggplot()+
  aes(x = Jinja_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Jinja_poly_3)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow - Density + Speed Model : Validation for Jinja road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))+
  scale_x_continuous(expand = c(0,0), limits = c(20,50), 
                    breaks = seq(20, 50, by = 5))+
  scale_y_continuous(expand = c(0,0), limits = c(20,50), 
                     breaks = seq(20, 50, by = 5))

  
```



## Combined


## Split data into training data set and validation data set
```{r}
## Split data into training data set and validation data set

n=dim(Combined_2)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Combined_2_train = Combined_2[id,]
Combined_2_validation = Combined_2[-id,]


```


### Model Summary three dimensions
```{r}

summary(Combined_2_unknown <- lm(data = Combined_2_train, 
                              formula = `Pedestrian Flow (peds/m/min)`  ~ 
                                0 + poly(`Pedestrian Density (Peds/m2) b`,2, raw = T) +
                                poly(1/(log10(`(Mean Speed m/min)`)),2, raw = T)))


coefficients(Combined_2_unknown)

```


### Model Validation Jinja three dimensions
```{r}

predict_Combined_2_poly_3 <- predict(Combined_2_unknown, 
                                newdata = Combined_2_validation)

ggplot()+
  aes(x = Combined_2_validation$`Pedestrian Flow (peds/m/min)`, 
      y = as.vector(predict_Combined_2_poly_3)) + 
  geom_point(color="red", position = "jitter")+
  geom_smooth(method = "lm", se = TRUE, color="black", 
              formula = y~x, size=0.5)+
  labs(title="Flow - Density + Speed Model : Validation for Combined road", 
       x="Actual Pedestrian Flow  peds/m/min", 
       y="Predicted Pedestrian Flow peds/m/min")+
  theme(plot.title = element_text(hjust = 0.5, size = 10))+
  theme(axis.title.y = element_text(size = 8))+
  theme(axis.title.x = element_text(size = 8))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"))

  
```



# PRINCIPAL COMPONENT ANALYSIS

## Enter Data
```{r}

Bombo_Clean = read_excel("Speed Density Flow.xlsx", 
                         sheet = "Bombo-Clean")

Entebbe_Clean = read_excel("Speed Density Flow.xlsx", 
                           sheet = "Entebbe-Clean")

Hoima_Clean = read_excel("Speed Density Flow.xlsx", 
                         sheet = "Hoima-Clean")


Jinja_Clean = read_excel("Speed Density Flow.xlsx", 
                         sheet = "Jinja-Clean")

Combined_3 = rbind(Bombo_Clean,Entebbe_Clean,Hoima_Clean,Jinja_Clean)
Combined_3[,12] <- Combined_3[,12]*10

```


## Split data into training data set and validation data set

```{r}

n=dim(Bombo)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Bombo_train = Bombo[id,]
Bombo_validation = Bombo[-id,]


n=dim(Entebbe)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Entebbe_train = Entebbe[id,]
Entebbe_validation = Entebbe[-id,]


n=dim(Hoima)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Hoima_train = Hoima[id,]
Hoima_validation = Hoima[-id,]


n=dim(Jinja)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Jinja_train = Jinja[id,]
Jinja_validation = Jinja[-id,]


n=dim(Combined_3)[1]
set.seed(12345)
id=sample(1:n, floor(n*0.8))
Combined_3_train = Combined_3[id,]
Combined_3_validation = Combined_3[-id,]

```


## Creating many new variables to be reduced by principal component analysis

```{r}

new_variables = cbind.data.frame(
          flow = Bombo_train$`Pedestrian Flow (peds/m/min)`,
          mean_speed = Bombo_train$`(Mean Speed m/min)`,
          mean_speed_2 = Bombo_train$`(Mean Speed m/min)`^2,
          mean_speed_3 = Bombo_train$`(Mean Speed m/min)`^3,
          log_speed = log(Bombo_train$`(Mean Speed m/min)`),
          inv_speed = Bombo_train$`(Mean Speed m/min)`^(-1),
          inv_log_speed = log(Bombo_train$`(Mean Speed m/min)`)^(-1),
          density = Bombo_train$`Pedestrian Density (Peds/m2) b`,
          density_2 = Bombo_train$`Pedestrian Density (Peds/m2) b`,
          density_3 = Bombo_train$`Pedestrian Density (Peds/m2) b`^3,
          log_density = log(Bombo_train$`Pedestrian Flow (peds/m/min)`),
          inv_density = Bombo_train$`Pedestrian Flow (peds/m/min)`^(-1),
          inv_log_density = log(Bombo_train$`Pedestrian Density (Peds/m2) b`)^(-1)
          )


head(new_variables, 10)


new_variables.pca <- prcomp(new_variables[,-1], center = TRUE,scale. = TRUE)

summary(new_variables.pca)

# str(new_variables.pca)

# library(devtools)
# install_github("vqv/ggbiplot")

```


## Sample of Principle Components
```{r}
head(new_variables.pca$x[,1:2])

head(new_variables.pca$rotation[, 1:2])

```



## Visualising principal component analysis

```{r}
library(ggbiplot)
ggbiplot(new_variables.pca)

```


## Applying principal component analysis

```{r}

new_variables.pca <- prcomp(new_variables[,-1], center = TRUE,scale. = TRUE)

summary(new_variables.pca)

str(new_variables.pca)

```

## Using PCA1 and PCA2 as variables in regression

```{r}

summary(flow_pca_prediction <- 
          lm(formula = new_variables$flow  ~ 
               new_variables.pca$x[,1] + 
               new_variables.pca$x[,2]))


summary(flow_pca_prediction <- 
          lm(formula = new_variables$flow  ~ 
               0 + poly(new_variables.pca$x[,1],2, raw = T) +
               poly(new_variables.pca$x[,2],2, raw = T) ))

summary(flow_pca_prediction <- 
          lm(formula = new_variables$flow  ~ 
               poly(new_variables.pca$x[,1],2, raw = T) +
               poly(new_variables.pca$x[,2],2, raw = T) ))

```


